---
title: "Assignment two markdown"
output:
  html_document: default
  word_document: default
---

## Packages used include;
- readxl
- tidyverse
- ggplot2
- ggpubr
- dplyr

```{r, echo= FALSE}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(dplyr)

```
# Importing data

```{r, include = FALSE}
work <- read_excel(("D:/R/Assignment2/Assignment2_dataset.xlsx"))
View(work)
```
# **Question 1**
### removing the missing values

```{r}
work <- na.omit(work)

```

# **Question 2**

### Show the relationship between the prices and perception change.

### Approach
- define the perception values.
- compute the percentage of each perception in response to price.

```{r, echo =  FALSE}
postive  <- ((length(c(work$P[work$P == "Positive"]))/length(c(work$P))))*100
negative <- ((length(c(work$P[work$P == "Negative"]))/length(c(work$P))))*100
partial_postive <- ((length(c(work$P[work$P == "SP"]))/length(c(work$P))))*100
partial_negative <- ((length(c(work$P[work$P == "SN"]))/length(c(work$P))))*100
Nuetral <- ((length(c(work$P[work$P == "NR"]))/length(c(work$P))))*100
```

## Results as percentages of all perceptions.

```{r}
postive 
negative
partial_postive
partial_negative
Nuetral
```

## Displaying the covariation between the perception and price

```{r}
ggplot(data = work, mapping = aes(x = price)) +
  geom_freqpoly(mapping = aes(color = P), binwidth = 500)
```

# **Question 3**

## Compare the perception change and diamond quality

```{r, echo=FALSE}
PC <- work$PC
cut <- work$cut
ggplot(data = work) +
  geom_count(mapping = aes(x = cut, y = PC))
```

# **Question 4**
## Generating a boxplot of all the variables
    
```{r, echo=FALSE}  
boxplot(work$price, work$carat, work$depth, work$x, work$y, names=c("price", "carat", "depth", "x", "y"), horizontal=TRUE, col="green", main="Boxplot of the variables of the dataset", xlab="Values", ylab="Variables")
```

# **Question 5**
## Generating induvidual boxplots of the variables

```{r, echo=FALSE}
boxplot(work$price, main="Boxplot of the price variable",col="green")
boxplot(work$carat, main="Boxplot of the carat variable",col="green")
boxplot(work$depth, main="Boxplot of the depth variable",col="green")
boxplot(work$x, main="Boxplot of the x variable",col="green")
boxplot(work$y, main="Boxplot of the y variable",col="green")
```

## A box plot that labels out the outliers
### Using the interquartile range to identify the outliers

### Approach
- Compute the interquartile range.
- Compute the upper and lower limits.
- Identify the outliers.

```{r,include=FALSE}
IQRprices <- IQR(work$price)
IQRprices

#For price.
quartileprice <- quantile(work$price, probs=c(.25, .75), na.rm = FALSE)
quartileprice
#calculate the upper and lower limits
pricelower <- quartileprice[1] - 1.5*IQRprices
priceupper <- quartileprice[2] + 1.5*IQRprices
pricelower
priceupper
#filter out the outliers for price
price_no_outliers <- subset(work, work$price > priceupper | work$price < pricelower)
price_no_outliers

## save the new dataframe with no outliers
new_work <- work[work$price < priceupper & work$price > pricelower,]

#For price/carat (x).
IQRx <- IQR(work$x)
IQRx
quartile_x <- quantile(work$x, probs=c(.25, .75), na.rm = FALSE)
quartile_x
#calculate the upper and lower limits
xlower <- quartileprice[1] - 1.5*IQRprices
xupper <- quartileprice[2] + 1.5*IQRprices
xlower
xupper
#filter out the outliers for x
x_no_outliers <- subset(work, work$x > xupper | work$x < xlower)
x_no_outliers
work_x <- work[work$x < priceupper & work$x > pricelower,]

IQRy <- IQR(work$y)
IQRy
#calculate the quartiles
quartilesy <- quantile(work$y, probs=c(.25, .75), na.rm = FALSE)
quartilesy
#calculate the upper and lower limits
Lowery <- quartilesy[1] - 1.5*IQRy
Uppery <- quartilesy[2] + 1.5*IQRy
Lowery
Uppery
#filter the outliers
no_outliersy <- subset(work, work$y > Uppery | work$y < Lowery)
no_outliersy
#box plot of the data without outliers

new_work <- work[work$y < Uppery & work$y > Lowery,]

```

# Plots without outliers

```{r,echo=FALSE}
boxplot(price_no_outliers$price, main="Boxplot of the price variable",col="blue")
boxplot(x_no_outliers$x , main="Boxplot of the x variable",col="blue")
boxplot(no_outliersy$y, main="Boxplot of the y variable",col="blue")
```

# **Question 6**
## creating a new csv file with the outliers removed

```{r}
write.csv(new_work, file="D:/R/Assignment2_MugangaCharles.csv")

```

# **Question 7**
## Display the relationship between one qualitative variable and one finite variable in the dataset
- the selected variables are cut and price.

```{r,echo=FALSE}
ggplot(new_work, aes(x=cut, y=price)) + geom_bar(stat="identity", fill="brown") + labs(title="Relationship between the cut and the price", x="Cut", y="Price")

```

# **Question 8**

## Compute the variance between three groups; diamond carat, perception change and price

### Approach
- Comparing mean, median and mode
- Calculating the mean and median of the variable "carat" in the diamonds dataset
```{r, echo=FALSE}

mean(new_work$carat,  na.rm = TRUE)
median(new_work$carat,  na.rm = TRUE)

```
## results 
- Mean =0.72, Median = 0.7.
- The performance of carat is positively skewed.

# **Question 9**
### Approach
- Compute the variance between three groups; diamond carat, perception change and price

- the groups are carat, price and perception change
- the variables are carat, PC and price
- The null hypothesis is that the groups have the same variance
- The alternative hypothesis is that the groups have different variances

#### Steps
- view the groups
```{r,echo=FALSE}
groups <- new_work[,c("carat", "PC", "price")] 
groups
```
- Generate the random sample of the data.
```{r,include=FALSE}
#generate ramdom sample of the data
sample <- sample_n(groups, 100)
sample
#the factors are the groups
#show the levels
#use factor function to define the groups
x <- factor(sample$PC)
x
levels(x)
#show the summary of the data by the groups
x <- ordered(x, levels=c("Negative","Positive","Positve","NR","SN","SP"))# the order of the levels
group_by(groups, PC) %>% 
  summarise(
    mean=mean(price), 
    sd=sd(price), n=n())
```

# Plots

```{r, echo=FALSE}
ggline(groups, x = "PC", y = "price", 
        add = c("mean_se"),
        order = c("Negative","Positive","Positve","NR","SN","SP"),
        ylab = "Price", xlab="Perception Change", 
        title = "Line plot of the groups")

ggline(groups, x = "PC", y = "carat", 
        add = c("mean_se"),
        order = c("Negative","Positive","Positve","NR","SN","SP"),
        ylab = "Carat", xlab="Perception Change", 
        title = "Line plot of the groups")
```

## Computng the variance betwen the groups
    
```{r, echo=FALSE}
anova <- aov(price ~ PC, data = groups)
summary(anova)
```

## Commenting on the results
- The p-value is 0.954 which is greater than 0.05 and therefore we fail to reject the null hypothesis therefore statistically not significant
- The posthoc test used is the Tukey HSD test
- The null hypothesis is that the groups have the same variance
- The alternative hypothesis is that the groups have different variances by aleast one group having a variance not equal to the others groups

# **Question 9(b)**

```{r}
TukeyHSD(anova, conf.level = .95)
plot(TukeyHSD(anova, conf.level = .95))
```

- The Tukey HSD test shows that the groups have different variances by aleast one group having a variance not equal to the others groups.